---
title: "Untitled"
author: "Rudolf Cesaretti"
date: "2023-03-07"
output: html_document
---


```{r, setup, include=FALSE,echo=FALSE, message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=75),tidy=TRUE)
#
rm(list = ls())
```



# Setup 

All of the data and scripts are downloadable from the [new ASU SettlementPersist2022 github repository](https://https://github.com/rcesaret/ASUSettlementPersist2022), which can be downloaded locally as a .zip folder or cloned to your own account.

Either way, once you have done so, you will need to modify the working directory (setwd("C:/...)") path and "dir" variables in the code chunk below to match the repository location on your computer.

```{r, label='Set Local Directory Location', message=FALSE,warning=FALSE}

wd <- list()

#SET YOUR LOCAL DIRECTORY LOCATION HERE:
#wd$dir <- "C:/Users/rcesaret/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"
wd$dir <- "C:/Users/TJ McMote/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"

wd$analysis <- paste0(wd$dir,"analysis/")
wd$data_r <- paste0(wd$dir,"data-raw/")
wd$data_p <- paste0(wd$dir,"data-processed/")
wd$data_f <- paste0(wd$dir,"data-final-outputs/")
wd$figs <- paste0(wd$dir,"figures/")
wd$funcs <- paste0(wd$dir,"functions/")

```


## Load R Packages and Custom Functions

```{r, label='Load Libraries', message=FALSE,warning=FALSE}
# Package names
packages <- c("rgdal", "sp", "sf", "GISTools", "lwgeom", "tidyverse", "tidyr", 
              "data.table", "zoo", "scales", "sjmisc", "REAT")
              
              #, "data.table",  "mgcv","igraph", "ggrepel","ggridges", "movecost",  "datplot", "scales",

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(packages,installed_packages)

#Read in custom R functions located in the wd$funcs directory folder
#FUNCS <- list("splitByAttributes.R", "RS_Acoef.R", "PopAreaResids.R", "LQ.R", "GiniSp.R")
#invisible(lapply(FUNCS, function(x) source(paste0(wd$funcs,x))))
#rm(FUNCS)

```


```{r, label='Import Data', message=FALSE,warning=FALSE}

# Read-in the data
MH_Settlements <- read.csv(paste0(wd$data_r,"MH_Settlements.csv"))
MH_Occu <- read.csv(paste0(wd$data_r,"MH_Occu.csv"))
MH_Attr <- read.csv(paste0(wd$data_r,"MH_Attr.csv"), na.strings	= "")


#MH_Poly <- st_read(paste0(wd$data_r,"MH_Poly.gpkg")) %>% rename(AggZone_ID = id, AggZone = Name)
```


```{r}
MH_Agg = MH_Settlements %>% select( -East, -North, -ORD, -AggZoneNumConserv, -ConservAgg_East, -ConservAgg_North, -AggZone_fid) %>% left_join(MH_Occu, by = "Name") %>% group_by(AggZone_ID, AggZone) %>% summarise_if(is.numeric, sum) %>% ungroup() %>% select(-ORD)

MH_Agg_long <- MH_Agg %>% pivot_longer(cols = OldFarm_T:CantSeeIfOccu_U,
  names_to = "Variable", values_to = "Count") %>% filter(Count != 0) %>% 
  mutate(presence = ifelse(Count > 0, 1, 0)) %>% 
  left_join(MH_Attr, by="Variable")

MH_Occu_long <- MH_Occu %>% pivot_longer(cols = OldFarm_T:CantSeeIfOccu_U,
  names_to = "Variable", values_to = "Count") %>% filter(Count != 0) %>% 
  left_join(MH_Attr, by="Variable")
#MH_Occu_long_Totals <- MH_Occu_long %>% filter(Status == "Total")

#MH_Occu_long_NoTotals <- MH_Occu_long %>% filter(Status != "Total")

```




# Calculate Macro Demog/Econ/Class Counts and Percents


```{r}
## Total CNpicto Pop Head Count (including attached/dependent noblewomen)

TotalCount = MH_Agg_long %>% filter(Demog != "Runaway") %>% filter(Demog != "Dead") %>% 
  group_by(AggZone, Suffix2) %>% summarise(TotalCount = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(TotalCount)) %>% ungroup() %>% 
  mutate(PctTotalCount = TotalCount / Totals) 

PctTotalCount <- TotalCount %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctTotalCount_", values_from = PctTotalCount, values_fill = 0) %>% 
  select(-PctTotalCount_T)

TotalCount <- TotalCount %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "TotalCount_", values_from = TotalCount, values_fill = 0) %>% 
  left_join(PctTotalCount, by = "AggZone")

rm(PctTotalCount)

## Heads of Household

TotalHH = MH_Agg_long %>% filter(Demog != "Runaway") %>% filter(Demog != "Dead") %>% 
  filter(OccuSpecInclude != "DemogExcl" & Demog != "FemaleNoble") %>%
  group_by(AggZone, Suffix2) %>% summarise(TotalHH = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(TotalHH)) %>% ungroup() %>% 
  mutate(PctTotalHH = TotalHH / Totals) 

PctTotalHH <- TotalHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctTotalHH_", values_from = PctTotalHH, values_fill = 0) %>% 
  select(-PctTotalHH_T)

TotalHH <- TotalHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "TotalHH_", values_from = TotalHH, values_fill = 0) %>% 
  left_join(PctTotalHH, by = "AggZone")

rm(PctTotalHH)

## Attached/Dependent Female Nobles

NobleAttachedFem = MH_Agg_long %>% filter(OccuSpecIncl == "DemogExcl" & Demog == "FemaleNoble") %>% 
  group_by(AggZone) %>% summarise(NobleAttachedFem = sum(Count))

## Commoner Households

CommonHH = MH_Agg_long %>% filter(Class == "Commoner") %>% filter(Demog != "Runaway") %>% filter(Demog != "Dead") %>% 
  group_by(AggZone, Suffix2) %>% summarise(CommonHH = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(CommonHH)) %>% ungroup() %>% 
  mutate(PctCommonHH = CommonHH / Totals) 

PctCommonHH <- CommonHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctCommonHH_", values_from = PctCommonHH, values_fill = 0) %>% 
  select(-PctCommonHH_T)

CommonHH <- CommonHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "CommonHH_", values_from = CommonHH, values_fill = 0) %>% 
  left_join(PctCommonHH, by = "AggZone")

rm(PctCommonHH)

## Recent Runaway heads of household

RunawayHH = MH_Agg_long %>% filter(Demog == "Runaway") %>% 
   group_by(AggZone) %>% summarise(RunawayHH = sum(Count))

## Recently Dead heads of household

DeadHH = MH_Agg_long %>% filter(Demog == "Dead") %>% 
   group_by(AggZone) %>% summarise(DeadHH = sum(Count))

## Tributary Households

TribHH = MH_Agg_long %>% filter(Trib == "Tributary") %>% 
  group_by(AggZone, Suffix2) %>% summarise(TribHH = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(TribHH)) %>% ungroup() %>% 
  mutate(PctTribHH = TribHH / Totals) 

PctTribHH <- TribHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctTribHH_", values_from = PctTribHH, values_fill = 0) %>% 
  select(-PctTribHH_T)

TribHH <- TribHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "TribHH_", values_from = TribHH, values_fill = 0) %>% 
  left_join(PctTribHH, by = "AggZone")

rm(PctTribHH)

## Exempt (Non-Tributary) Commoner Households

ExemptHH = MH_Agg_long %>% filter(Trib == "Exempt") %>% 
  group_by(AggZone, Suffix2) %>% summarise(ExemptHH = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(ExemptHH)) %>% ungroup() %>% 
  mutate(PctExemptHH = ExemptHH / Totals) 

PctExemptHH <- ExemptHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctExemptHH_", values_from = PctExemptHH, values_fill = 0) %>% 
  select(-PctExemptHH_T)

ExemptHH <- ExemptHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "ExemptHH_", values_from = ExemptHH, values_fill = 0) %>% 
  left_join(PctExemptHH, by = "AggZone")

rm(PctExemptHH)

## Married Elderly/Older Men (Non-Tributary) Commoner Households

OldHH = MH_Agg_long %>% filter(Demog == "Elderly") %>% 
  group_by(AggZone, Suffix2) %>% summarise(OldHH = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(OldHH)) %>% ungroup() %>% 
  mutate(PctOldHH = OldHH / Totals) 

PctOldHH <- OldHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctOldHH_", values_from = PctOldHH, values_fill = 0) %>% 
  select(-PctOldHH_T)

OldHH <- OldHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "OldHH_", values_from = OldHH, values_fill = 0) %>% 
  left_join(PctOldHH, by = "AggZone")

rm(PctOldHH)

## Widower (Non-Tributary) Commoner Households

WidowerHH = MH_Agg_long %>% filter(Demog == "Widower") %>% 
  group_by(AggZone, Suffix2) %>% summarise(WidowerHH = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(WidowerHH)) %>% ungroup() %>% 
  mutate(PctWidowerHH = WidowerHH / Totals) 

PctWidowerHH <- WidowerHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctWidowerHH_", values_from = PctWidowerHH, values_fill = 0) %>% 
  select(-PctWidowerHH_T)

WidowerHH <- WidowerHH %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "WidowerHH_", values_from = WidowerHH, values_fill = 0) %>% 
  left_join(PctWidowerHH, by = "AggZone")

rm(PctWidowerHH)

## Widow (Non-Tributary) Commoner Heads of Household

WidowHH = MH_Agg_long %>% filter(Demog == "Widow") %>% 
  group_by(AggZone, Suffix2) %>% summarise(WidowHH = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(WidowHH)) %>% ungroup() %>% 
  mutate(PctWidowHH = WidowHH / Totals) %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "WidowHH_", values_from = WidowHH, values_fill = 0)

## Sick/Infirm/Disabled (Non-Tributary) Commoner Heads of Household

SickHH = MH_Agg_long %>% filter(Demog == "Sick") %>% 
  group_by(AggZone, Suffix2) %>% summarise(SickHH = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(SickHH)) %>% ungroup() %>% 
  mutate(PctSickHH = SickHH / Totals) %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "SickHH_", values_from = SickHH, values_fill = 0)

## Tributary Occupational Specialists

TribSpecialists = MH_Agg_long %>% filter(OccuSpecInclude == 1) %>% 
  group_by(AggZone, Suffix2) %>% summarise(TribSpecialists = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(TribSpecialists)) %>% ungroup() %>% 
  mutate(PctTribSpecialists = TribSpecialists / Totals) 

PctTribSpecialists <- TribSpecialists %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctTribSpecialists_", values_from = PctTribSpecialists, values_fill = 0) %>% 
  select(-PctTribSpecialists_T)

TribSpecialists <- TribSpecialists %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "TribSpecialists_", values_from = TribSpecialists, values_fill = 0) %>% 
  left_join(PctTribSpecialists, by = "AggZone")

rm(PctTribSpecialists)

## Tributary + Elderly Occupational Specialists

SpecialistsO = MH_Agg_long %>% filter(OccuSpecInclude_O == 1) %>% 
  group_by(AggZone, Suffix2) %>% summarise(SpecialistsO = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(SpecialistsO)) %>% ungroup() %>% 
  mutate(PctSpecialistsO = SpecialistsO / Totals) 

PctSpecialistsO <- SpecialistsO %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctSpecialistsO_", values_from = PctSpecialistsO, values_fill = 0) %>% 
  select(-PctSpecialistsO_T)

SpecialistsO <- SpecialistsO %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "SpecialistsO_", values_from = SpecialistsO, values_fill = 0) %>% 
  left_join(PctSpecialistsO, by = "AggZone")

rm(PctSpecialistsO)

## Tributary + Noble Occupational Specialists

SpecialistsN = MH_Agg_long %>% filter(OccuSpecInclude_N == 1) %>% 
  group_by(AggZone, Suffix2) %>% summarise(SpecialistsN = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(SpecialistsN)) %>% ungroup() %>% 
  mutate(PctSpecialistsN = SpecialistsN / Totals) 

PctSpecialistsN <- SpecialistsN %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctSpecialistsN_", values_from = PctSpecialistsN, values_fill = 0) %>% 
  select(-PctSpecialistsN_T)

SpecialistsN <- SpecialistsN %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "SpecialistsN_", values_from = SpecialistsN, values_fill = 0) %>% 
  left_join(PctSpecialistsN, by = "AggZone")

rm(PctSpecialistsN)

## Tributary Occupational Specialists + Farmers 

SpecialistsF = MH_Agg_long %>% filter(OccuSpecInclude_F == 1) %>% 
  group_by(AggZone, Suffix2) %>% summarise(SpecialistsF = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(SpecialistsF)) %>% ungroup() %>% 
  mutate(PctSpecialistsF = SpecialistsF / Totals) 

PctSpecialistsF <- SpecialistsF %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctSpecialistsF_", values_from = PctSpecialistsF, values_fill = 0) %>% 
  select(-PctSpecialistsF_T)

SpecialistsF <- SpecialistsF %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "SpecialistsF_", values_from = SpecialistsF, values_fill = 0) %>% 
  left_join(PctSpecialistsF, by = "AggZone")

rm(PctSpecialistsF)

## Tributary Occupational Specialists + Elderly + Nobles 

SpecialistsNO = MH_Agg_long %>% filter(OccuSpecInclude_NO == 1) %>% 
  group_by(AggZone, Suffix2) %>% summarise(SpecialistsNO = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(SpecialistsNO)) %>% ungroup() %>% 
  mutate(PctSpecialistsNO = SpecialistsNO / Totals) 

PctSpecialistsNO <- SpecialistsNO %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctSpecialistsNO_", values_from = PctSpecialistsNO, values_fill = 0) %>% 
  select(-PctSpecialistsNO_T)

SpecialistsNO <- SpecialistsNO %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "SpecialistsNO_", values_from = SpecialistsNO, values_fill = 0) %>% 
  left_join(PctSpecialistsNO, by = "AggZone")

rm(PctSpecialistsNO)

## Tributary Occupational Specialists + Elderly + Farmers 

SpecialistsFO = MH_Agg_long %>% filter(OccuSpecInclude_FO == 1) %>% 
  group_by(AggZone, Suffix2) %>% summarise(SpecialistsFO = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(SpecialistsFO)) %>% ungroup() %>% 
  mutate(PctSpecialistsFO = SpecialistsFO / Totals) 

PctSpecialistsFO <- SpecialistsFO %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctSpecialistsFO_", values_from = PctSpecialistsFO, values_fill = 0) %>% 
  select(-PctSpecialistsFO_T)

SpecialistsFO <- SpecialistsFO %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "SpecialistsFO_", values_from = SpecialistsFO, values_fill = 0) %>% 
  left_join(PctSpecialistsFO, by = "AggZone")

rm(PctSpecialistsFO)

## Tributary Occupational Specialists + Elderly + Farmers 

SpecialistsNF = MH_Agg_long %>% filter(OccuSpecInclude_NF == 1) %>% 
  group_by(AggZone, Suffix2) %>% summarise(SpecialistsNF = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(SpecialistsNF)) %>% ungroup() %>% 
  mutate(PctSpecialistsNF = SpecialistsNF / Totals) 

PctSpecialistsNF <- SpecialistsNF %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctSpecialistsNF_", values_from = PctSpecialistsNF, values_fill = 0) %>% 
  select(-PctSpecialistsNF_T)

SpecialistsNF <- SpecialistsNF %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "SpecialistsNF_", values_from = SpecialistsNF, values_fill = 0) %>% 
  left_join(PctSpecialistsNF, by = "AggZone")

rm(PctSpecialistsNF)

## Tributary Occupational Specialists + Elderly + Farmers + Nobles

SpecialistsNFO = MH_Agg_long %>% filter(OccuSpecInclude_NFO == 1) %>% 
  group_by(AggZone, Suffix2) %>% summarise(SpecialistsNFO = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(SpecialistsNFO)) %>% ungroup() %>% 
  mutate(PctSpecialistsNFO = SpecialistsNFO / Totals) 

PctSpecialistsNFO <- SpecialistsNFO %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctSpecialistsNFO_", values_from = PctSpecialistsNFO, values_fill = 0) %>% 
  select(-PctSpecialistsNFO_T)

SpecialistsNFO <- SpecialistsNFO %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "SpecialistsNFO_", values_from = SpecialistsNFO, values_fill = 0) %>% 
  left_join(PctSpecialistsNFO, by = "AggZone")

rm(PctSpecialistsNFO)

## Old / Elderly Married Men (Non-Tributary) Occupational Specialists

OldSpecialists = MH_Agg_long %>% filter(OccuSpecIncl == "ExemptSpec") %>% 
  group_by(AggZone, Suffix2) %>% summarise(OldSpecialists = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(OldSpecialists)) %>% ungroup() %>% 
  mutate(PctOldSpecialists = OldSpecialists / Totals) 

PctOldSpecialists <- OldSpecialists %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctOldSpecialists_", values_from = PctOldSpecialists, values_fill = 0) 

OldSpecialists <- OldSpecialists %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "OldSpecialists_", values_from = OldSpecialists, values_fill = 0) %>% 
  left_join(PctOldSpecialists, by = "AggZone") %>% 
  mutate(OldSpecialists_T = OldSpecialists_F + OldSpecialists_R + OldSpecialists_U)

rm(PctOldSpecialists)

## Tributary Farmer Households

TribFarmers = MH_Agg_long %>% filter(Occupation2 == "Farmer") %>% 
  group_by(AggZone, Suffix2) %>% summarise(TribFarmers = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(TribFarmers)) %>% ungroup() %>% 
  mutate(PctTribFarmers = TribFarmers / Totals) 

PctTribFarmers <- TribFarmers %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctTribFarmers_", values_from = PctTribFarmers, values_fill = 0) %>% 
  select(-PctTribFarmers_T)

TribFarmers <- TribFarmers %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "TribFarmers_", values_from = TribFarmers, values_fill = 0) %>% 
  left_join(PctTribFarmers, by = "AggZone")

rm(PctTribFarmers)

## Exempt / Non-Tributary Farmer Households

ExemptFarmers = MH_Agg_long %>% filter(Occupation == "Farmer" & Trib == "Exempt") %>% 
  group_by(AggZone, Suffix2) %>% summarise(ExemptFarmers = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(ExemptFarmers)) %>% ungroup() %>% 
  mutate(PctExemptFarmers = ExemptFarmers / Totals) 

PctExemptFarmers <- ExemptFarmers %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctExemptFarmers_", values_from = PctExemptFarmers, values_fill = 0) %>% 
  select(-PctExemptFarmers_T)

ExemptFarmers <- ExemptFarmers %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "ExemptFarmers_", values_from = ExemptFarmers, values_fill = 0) %>% 
  left_join(PctExemptFarmers, by = "AggZone")

rm(PctExemptFarmers)

## Total Farmer Households - including Tributaries + Exempt/Non-Tributaries

TotalFarmers = MH_Agg_long %>% filter(Occupation == "Farmer") %>% 
  group_by(AggZone, Suffix2) %>% summarise(TotalFarmers = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(TotalFarmers)) %>% ungroup() %>% 
  mutate(PctTotalFarmers = TotalFarmers / Totals) 

PctTotalFarmers <- TotalFarmers %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctTotalFarmers_", values_from = PctTotalFarmers, values_fill = 0) %>% 
  select(-PctTotalFarmers_T)

TotalFarmers <- TotalFarmers %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "TotalFarmers_", values_from = TotalFarmers, values_fill = 0) %>% 
  left_join(PctTotalFarmers, by = "AggZone")

rm(PctTotalFarmers)

## Total Farmer Households - including Tributaries + Exempt/Non-Tributaries + Widows

TotalFarmersW = MH_Agg_long %>% filter(Occupation == "Farmer" | Occupation == "Widow") %>% 
  group_by(AggZone, Suffix2) %>% summarise(TotalFarmersW = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone) %>% mutate(Totals = max(TotalFarmersW)) %>% ungroup() %>% 
  mutate(PctTotalFarmersW = TotalFarmersW / Totals) 

PctTotalFarmersW <- TotalFarmersW %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "PctTotalFarmersW_", values_from = PctTotalFarmersW, values_fill = 0) %>% 
  select(-PctTotalFarmersW_T)

TotalFarmersW <- TotalFarmersW %>% pivot_wider(id_cols = AggZone, names_from = Suffix2,
  names_prefix = "TotalFarmersW_", values_from = TotalFarmersW, values_fill = 0) %>% 
  left_join(PctTotalFarmersW, by = "AggZone")

rm(PctTotalFarmersW)


```


```{r}
x = TotalCount %>% left_join(TotalHH, by = "AggZone") %>% left_join(NobleAttachedFem, by = "AggZone") %>%
  left_join(CommonHH, by = "AggZone") %>% left_join(RunawayHH, by = "AggZone") %>% 
  left_join(DeadHH, by = "AggZone") %>% left_join(TribHH, by = "AggZone") %>% 
  left_join(ExemptHH, by = "AggZone") %>% left_join(OldHH, by = "AggZone") %>% 
  left_join(WidowerHH, by = "AggZone") %>% left_join(WidowHH, by = "AggZone") %>%
  left_join(SickHH, by = "AggZone") %>% left_join(TribSpecialists, by = "AggZone") %>%
  left_join(SpecialistsO, by = "AggZone") %>% left_join(SpecialistsN, by = "AggZone") %>%
  left_join(SpecialistsF, by = "AggZone") %>% left_join(SpecialistsNO, by = "AggZone") %>%
  left_join(SpecialistsFO, by = "AggZone") %>% left_join(SpecialistsNF, by = "AggZone") %>%
  left_join(SpecialistsNFO, by = "AggZone") %>% left_join(OldSpecialists, by = "AggZone") %>%
  left_join(TribFarmers, by = "AggZone") %>% left_join(ExemptFarmers, by = "AggZone") %>%
  left_join(TotalFarmers, by = "AggZone") %>% left_join(TotalFarmersW, by = "AggZone")
  
x[is.na(x)] <- 0
MH_Agg_Occu <- MH_Agg %>% select(AggZone_ID:AggZone, OldFarm_T:CantSeeIfOccu_U)
MH_Agg <- MH_Agg %>% select(AggZone_ID:CircleRenters) %>% left_join(x, by = "AggZone") 
```

```{r}
rm(x, TotalCount, TotalHH, NobleAttachedFem, CommonHH, RunawayHH, DeadHH, TribHH,
ExemptHH, OldHH, WidowerHH, WidowHH, SickHH, TribSpecialists, SpecialistsO,
SpecialistsN, SpecialistsF, SpecialistsNO, SpecialistsFO, SpecialistsNF,
SpecialistsNFO, OldSpecialists, TribFarmers, ExemptFarmers, TotalFarmers, TotalFarmersW)
```




# Calculate Status Percentages for Occupations

```{r}

AggOccuPercents = MH_Agg_Occu %>% select(AggZone, FieldHand_T:CantSeeIfOccu_U) %>%
  pivot_longer(cols = FieldHand_T:CantSeeIfOccu_U, names_to = "Variable", values_to = "Count") %>%
  #filter(Count != 0) %>% 
  left_join(MH_Attr, by="Variable") %>%
  group_by(AggZone, Occupation2) %>% 
  mutate(CountTotal = max(Count)) %>% ungroup() %>%  
  mutate(PctTotal = Count / CountTotal) %>% 
  filter(Status == "Renter" | Status == "Free" | Status == "Uncertain") %>%
  select(AggZone, Variable, PctTotal) %>% mutate_all(~replace(., is.na(.), 0)) %>%
  pivot_wider(id_cols = AggZone, names_from = Variable, names_prefix = "Pct", 
              values_from = PctTotal)
  
```


# 


# Calculate Sector Counts and Percentages



This new code works perfectly, but the MH_Attr dataframe has:
--Primary sector includes Old, widowers, widows
--nobles are included in Tertiary/Quaternary
--etc.

Perhaps 4 versions for each sector?
-nobles
-exempt
-nobles + exempt
-tributaries only (this DOES include farmers)

Jettison 3b and 4b? 4? 4 = Quaternart; 3b + 4b = PriSec, SecTert, etc.
--perhaps just get rid of 4 + 4b -- these are not very standard in economics or econ hist
--keep 3b
--keep subsector + industry
--NEED TO CREATE A NEW SECTOR CATEGORY THAT CONFORMS TO EMPIRICAL PATTERNS
    Spatial distribution, renters, part-time/low-intensity vs. higher-intersity/higher-specialization
    in terms of labor time, capital inputs, scale of production (i.e. single household? or need wider organization
    and dependent on wider division of labor FOR PRODUCTION of the good or service e.g. stoneworkers, featherworkers, etc.)

Also, do you want to get rid of the "Unk" (unknown) sector??

```{r}
## Sectors3

Sectors3 = MH_Agg_long %>% group_by(AggZone, Sectors3, Suffix2) %>%
  summarise(TotalCount = sum(Count)) %>% ungroup() %>% 
  group_by(AggZone, Sectors3) %>% mutate(Totals = max(TotalCount)) %>% ungroup() %>% 
  mutate(PctTotalCount = TotalCount / Totals) %>% filter(!is.na(Sectors3)) 

PctSectors3 <- Sectors3 %>% mutate(x = paste(Sectors3,"_",Suffix2)) %>% 
  filter(Suffix2 != "T") %>% select(-TotalCount, -Totals, -Sectors3, -Suffix2) %>% 
  pivot_wider(id_cols = AggZone, names_from = x, names_prefix = "PctSectors3_", 
  values_from = PctTotalCount)

Sectors3 <- Sectors3 %>% mutate(x = paste(Sectors3,"_",Suffix2)) %>% 
  select(-Sectors3, -Suffix2) %>% pivot_wider(id_cols = AggZone, names_from = x,
  names_prefix = "Sectors3_", values_from = TotalCount, values_fill = 0) %>% 
  left_join(PctSectors3, by = "AggZone")

rm(PctSectors3)
```





```{r}




Sect3_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Sectors3) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3)) %>% pivot_wider(id_cols=Name, names_from = Sectors3, names_prefix = "Sect3_Tot_", values_from = Count, values_fill = 0)

Sect3_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors3) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3)) %>% pivot_wider(id_cols=Name, names_from = Sectors3, names_prefix = "Sect3_Tot_NoPix_", values_from = Count, values_fill = 0)

Sect3_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Sectors3) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3)) %>% pivot_wider(id_cols=Name, names_from = Sectors3, names_prefix = "Sect3_Trib_", values_from = Count, values_fill = 0)

Sect3_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors3) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3)) %>% pivot_wider(id_cols=Name, names_from = Sectors3, names_prefix = "Sect3_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Sectors4

Sect4_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Sectors4) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4)) %>% pivot_wider(id_cols=Name, names_from = Sectors4, names_prefix = "Sect4_Tot_", values_from = Count, values_fill = 0)

Sect4_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors4) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4)) %>% pivot_wider(id_cols=Name, names_from = Sectors4, names_prefix = "Sect4_Tot_NoPix_", values_from = Count, values_fill = 0)

Sect4_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Sectors4) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4)) %>% pivot_wider(id_cols=Name, names_from = Sectors4, names_prefix = "Sect4_Trib_", values_from = Count, values_fill = 0)

Sect4_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors4) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4)) %>% pivot_wider(id_cols=Name, names_from = Sectors4, names_prefix = "Sect4_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Sectors3b

Sect3b_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Sectors3b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3b)) %>% pivot_wider(id_cols=Name, names_from = Sectors3b, names_prefix = "Sect3b_Tot_", values_from = Count, values_fill = 0)

Sect3b_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors3b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3b)) %>% pivot_wider(id_cols=Name, names_from = Sectors3b, names_prefix = "Sect3b_Tot_NoPix_", values_from = Count, values_fill = 0)

Sect3b_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Sectors3b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3b)) %>% pivot_wider(id_cols=Name, names_from = Sectors3b, names_prefix = "Sect3b_Trib_", values_from = Count, values_fill = 0)

Sect3b_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors3b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3b)) %>% pivot_wider(id_cols=Name, names_from = Sectors3b, names_prefix = "Sect3b_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Sectors4b

Sect4b_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Sectors4b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4b)) %>% pivot_wider(id_cols=Name, names_from = Sectors4b, names_prefix = "Sect4b_Tot_", values_from = Count, values_fill = 0)

Sect4b_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors4b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4b)) %>% pivot_wider(id_cols=Name, names_from = Sectors4b, names_prefix = "Sect4b_Tot_NoPix_", values_from = Count, values_fill = 0)

Sect4b_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Sectors4b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4b)) %>% pivot_wider(id_cols=Name, names_from = Sectors4b, names_prefix = "Sect4b_Trib_", values_from = Count, values_fill = 0)

Sect4b_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors4b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4b)) %>% pivot_wider(id_cols=Name, names_from = Sectors4b, names_prefix = "Sect4b_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Subsectors

SubSect_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Subsector) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Subsector)) %>% pivot_wider(id_cols=Name, names_from = Subsector, names_prefix = "SubSect_Tot_", values_from = Count, values_fill = 0)

SubSect_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Subsector) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Subsector)) %>% pivot_wider(id_cols=Name, names_from = Subsector, names_prefix = "SubSect_Tot_NoPix_", values_from = Count, values_fill = 0)

SubSect_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Subsector) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Subsector)) %>% pivot_wider(id_cols=Name, names_from = Subsector, names_prefix = "SubSect_Trib_", values_from = Count, values_fill = 0)

SubSect_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Subsector) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Subsector)) %>% pivot_wider(id_cols=Name, names_from = Subsector, names_prefix = "SubSect_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Industries

Industry_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Industry) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Industry)) %>% pivot_wider(id_cols=Name, names_from = Industry, names_prefix = "Industry_Tot_", values_from = Count, values_fill = 0)

Industry_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Industry) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Industry)) %>% pivot_wider(id_cols=Name, names_from = Industry, names_prefix = "Industry_Tot_NoPix_", values_from = Count, values_fill = 0)

Industry_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Industry) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Industry)) %>% pivot_wider(id_cols=Name, names_from = Industry, names_prefix = "Industry_Trib_", values_from = Count, values_fill = 0)

Industry_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Industry) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Industry)) %>% pivot_wider(id_cols=Name, names_from = Industry, names_prefix = "Industry_Trib_NoPix_", values_from = Count, values_fill = 0)


```







# Overall Occupation Numbers, Ubiquity, Concentration and Dispersion

--include nobles
--do this for North, Central, South regions

Table:
-number of HH employed in occupation
-% of the total occupation employment in each of 3-4 regions (surroundings?)
-% renters of each occupation by region (3-4) and total
-Gini index as metric of occupational spatial concentration by region and for total


```{r}

AggZones = MH_Settlements %>% select(AggZone_ID, AggZone, ORD, Name)

#MH_Agg_Occu = AggZones %>% left_join(MH_Occu, by = c("ORD", "Name")) %>% select(-ORD, -Name) %>% group_by(AggZone_ID, AggZone) %>% summarise_if(is.numeric, sum) %>% ungroup 

MH_Agg_Occu_Long = MH_Agg_Occu %>% pivot_longer(cols = OldFarm_T:CantSeeIfOccu_U,
  names_to = "Variable", values_to = "Count") %>% filter(Count != 0) %>% 
  left_join(MH_Attr, by="Variable") 

MH_Agg_Occu_Long = MH_Agg_Occu_Long %>% mutate(presence = ifelse(Count > 0, 1, 0)) %>% 
  group_by(Variable) %>% mutate(occ.specialists = sum(Count, na.rm = T), 
                                ubiquity = sum(presence, na.rm = T)) %>% ungroup 

y = MH_Agg_Occu_Long %>% group_by(Variable, occ.specialists, ubiquity) %>% 
  summarise() %>% ungroup() %>% rename(Ubiquity_Agg = ubiquity)
  
MH_Occu_Long2 = MH_Occu_long %>% left_join(AggZones, by = c("ORD", "Name")) %>% 
  mutate(presence = ifelse(Count > 0, 1, 0)) %>% group_by(Variable) %>% 
  mutate(occ.specialists = sum(Count, na.rm = T), 
         ubiquity = sum(presence, na.rm = T)) %>% ungroup 

z = MH_Occu_Long2 %>% group_by(Variable, occ.specialists, ubiquity) %>% 
  summarise() %>% ungroup() %>% rename(Ubiquity_Barrios = ubiquity)

x = y %>% full_join(z, by = c("Variable", "occ.specialists"))

MH_Attr2 = MH_Attr %>% left_join(x, by = "Variable") %>% 
  rename(TotalCountOccu = occ.specialists) %>%
  mutate(TotalCountOccu = coalesce(TotalCountOccu, 0),
         Ubiquity_Agg = coalesce(Ubiquity_Agg, 0),
         Ubiquity_Barrios = coalesce(Ubiquity_Barrios, 0))

y = disp(MH_Agg_Occu[c(3:277)])
y = as.data.frame(t(as.matrix(y))) 
y = y %>% select(-Theil, -Dalton)
colnames(y) <- paste(colnames(y), "Agg", sep = "_")
y$Variable = rownames(y)

x = disp(MH_Occu[c(3:277)])
x = as.data.frame(t(as.matrix(x))) 
x = x %>% select(-Theil, -Dalton)
colnames(x) <- paste(colnames(x), "Bar", sep = "_")
x$Variable = rownames(x)

z = x %>% left_join(y, by="Variable")
names(z) <- c("Gini_Bar", "GiniNorm_Bar", "HHI_Bar", "HHINorm_Bar", "HHIeq_Bar", "HooverNon-Wt_Bar", "CoulterNon-Wt_Bar", "Atkinson_Bar", "SD_Bar", "CV_Bar", "CVNorm_Bar", "Variable", "Gini_Agg", "GiniNorm_Agg", "HHI_Agg", "HHINorm_Agg", "HHIeq_Agg", "HooverNon-Wt_Agg", "CoulterNon-Wt_Agg","Atkinson_Agg", "SD_Agg", "CV_Agg", "CVNprm_Agg")
MH_Attr2 = MH_Attr2 %>% left_join(z, by = "Variable")

MH_Attr2 = MH_Attr2 %>% select(-GiniNorm_Bar, -HHINorm_Bar, -GiniNorm_Agg, -HHINorm_Agg, -CV_Bar, -CV_Agg)

PctRenter = MH_Attr2 %>% group_by(Occupation2) %>% mutate(OccuTotal = max(TotalCountOccu)) %>% 
  ungroup() %>% mutate(PctRenter = TotalCountOccu / OccuTotal) %>% filter(Status == "Renter") %>% 
  select(Occupation2, PctRenter, TotalCountOccu) %>% rename(Renter = TotalCountOccu) %>% 
  group_by(Occupation2) %>% summarise_all(sum)
  
PctFree = MH_Attr2 %>% group_by(Occupation2) %>% mutate(OccuTotal = max(TotalCountOccu)) %>% 
  ungroup() %>% mutate(PctFree = TotalCountOccu / OccuTotal) %>% filter(Status == "Free") %>% 
  select(Occupation2, PctFree, TotalCountOccu) %>% rename(Free = TotalCountOccu) %>% 
  group_by(Occupation2) %>% summarise_all(sum)

x = PctRenter %>% left_join(PctFree, by="Occupation2") %>% mutate(RF_Ratio = Renter / Free,
                                                                  FR_Ratio = Free / Renter)

tot = MH_Attr2 %>% filter(Status == "Total") %>% left_join(x, by="Occupation2") %>% 
  filter(OccuSpecInclude == 1) #%>% select(-Class:ExemptSpecialists, -English:Nahuatl)

write.csv(tot, paste0(wd$data_p,"MH_Attr2.csv"))

rm(x,y,z,tot)

```




# Balassa/LQ for Occupations and Sectors

--It's BAlassa, NOT BElassa!!!!

--Balassa_pop == Balassa!! Total employment == total households or total tributaries
--Total Employment == economically active population
--Total Employment =/= Total specialists! Farmers are employed in farming!!

Calculate = 
--Balassa

--Additive Balassa = Hoen and Oosterhaven
-- regional specialisation coefficient  = Hoen and Oosterhaven

--  A modification of the Gini coefficient is used for the spatial concentration of industries as well as regional specialization (e.g. Ceapraz 2008). As we can see in the calculation of Ri and Cj, respectively, the spatial Gini coefficient is based on the LQ.

--Cluster Index by Litzenberger and Sternberg --> needs number of FIRMS
----FIRMS == groupings of specialists 






















