---
title: "Huexotzinco-Tlaxcala C16, Script #X:"
subtitle: "Topography, Environment, Agriculture and Subsistence"
author: "Rudolf Cesaretti"
date: "Last run on `r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
    number_sections: true
bibliography: References.bib
csl: apa.csl
link-citations: yes
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

```{r, setup, include=FALSE,echo=FALSE, message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=75),tidy=TRUE)
#
rm(list = ls())
```

I do four things in this R markdown document: 
Topographic/environmental metrics

  1. Calculate catchment area variables, including
      + Topographic/environmental metrics
      + 
  4. Reorganize the data and export for Script #6
  
  
# Setup 

All of the data and scripts are downloadable from the [new ASU SettlementPersist2022 github repository](https://https://github.com/rcesaret/ASUSettlementPersist2022), which can be downloaded locally as a .zip folder or cloned to your own account.

Either way, once you have done so, you will need to modify the working directory (setwd("C:/...)") path and "dir" variables in the code chunk below to match the repository location on your computer.

```{r, label='Set Local Directory Location', message=FALSE,warning=FALSE}

wd <- list()

#SET YOUR LOCAL DIRECTORY LOCATION HERE:
#wd$dir <- "C:/Users/rcesaret/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"
wd$dir <- "C:/Users/TJ McMote/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"

wd$analysis <- paste0(wd$dir,"analysis/")
wd$data_r <- paste0(wd$dir,"data-raw/")
wd$data_p <- paste0(wd$dir,"data-processed/")
wd$data_f <- paste0(wd$dir,"data-final-outputs/")
wd$figs <- paste0(wd$dir,"figures/")
wd$funcs <- paste0(wd$dir,"functions/")

```


## Load R Packages and Custom Functions

```{r, label='Load Libraries', message=FALSE,warning=FALSE}
# Package names
packages <- c("rgdal", "rgeos", "sp", "sf", "GISTools", "raster", "Matrix", "terra",
              "gdistance", "lwgeom", "tidyverse", "tidyr", "stars", "dismo", 
              "purrr", "spatialEco", "whitebox", "classInt","ggnewscale")#, "data.table", "zoo", "era", "JOPS", "mgcv","igraph",  "ggrepel","ggridges", "movecost",  "datplot", "scales",

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(packages,installed_packages)

#Read in custom R functions located in the wd$funcs directory folder
FUNCS <- list("RescaleSpatRast.R")#"splitByAttributes.R", 
invisible(lapply(FUNCS, function(x) source(paste0(wd$funcs,x))))
rm(FUNCS)

```


# Import and Reorganize Data

Data we are importing:

  1. 
  2. 
  3. 
  4. 

# Elevation and Slope

```{r, label='Import Data', message=FALSE,warning=FALSE}

## Import Smoothed 90 meter DEM
DEM_r <- terra::rast(raster(paste0(wd$data_r, "SRTM_30m_DEM_PueblaTlaxcala2_Smooth3.tif")))
DEM = read_stars(paste0(wd$data_r, "SRTM_30m_DEM_PueblaTlaxcala2_Smooth3.tif"))

#Hillshade = read_stars(paste0(wd$data_r, "Hillshade_PueblaTlaxcala2.tif"))

#HuexArable = read_stars(paste0(wd$data_r, "HuexArable5.tif"))

#DEM_r <- terra::rast(raster("D:/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/GIS_Data/DEM_CEM_v3.0_15m_CMex.tif"))
```
"D:/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/GIS_Data/Slope_CEM_v3.0_15m_CMex.tif"
"D:/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/GIS_Data/DEM_CEM_v3.0_15m_CMex.tif"



CHELSA_bio10_MeanPrecip.tif
CHELSA_bio10_meanTemp.tif
CHELSA_gsl_GrowingSeasonLength.tif
CHELSA_gsp_GrowingSeasonPrecipSum.tif
CHELSA_gst_GrowingSeasonMeanTemp.tif
CHELSA_NPP.tif

SoilGrids250_BulkDensity.tif
SoilGrids250_Clay.tif
SoilGrids250_CoarseFragments.tif
SoilGrids250_Nitrogen.tif
SoilGrids250_OrganicCarbonDensity.tif
SoilGrids250_Sand.tif
SoilGrids250_Silt.tif
SoilGrids250_SoilOrganaicCarbon.tif
SoilGrids250_SoilOrganaicCarbonStock.tif
SoilGrids250_WRBSoilType.tif

x = c("CHELSA_bio10_MeanPrecip.tif", "CHELSA_bio10_meanTemp.tif")
file_list = as.list(paste0(wd$data_r, x))
(CHELSA = read_stars(file_list, quiet = TRUE))



## Modern Avg Yeild Data

```{r, label='Import and Clean Modern Avg Yeild Data', message=FALSE,warning=FALSE}

#Import polygons of cultivated area by municipio and agriculture type (irrigation vs temporal)
HuexModernCultivArea_Municip_AGType <- st_read(paste0(wd$data_r,"HuexModernCultivArea_Municip_AGType.gpkg")) %>% rename(Municipio = NOM_MUN)

# these are the target municipios with polygons in the Puebla-Tlaxcala Valley we will analyze 
mun = unique(HuexModernCultivArea_Municip_AGType$Municipio)

#Import mexican municipal agricultural production stats
x = read.csv(paste0(wd$data_r,"Cierre_agricola_mun_2003_2021.csv"))

# clean and summarize the data
x = x %>% filter(Nomcultivo == "Maiz grano") %>% filter(Nomestado == "Puebla" | Nomestado == "Tlaxcala") %>% filter(Nomcicloproductivo == "Primavera-Verano") %>% filter(Nommunicipio %in% mun) %>% mutate(Rendimiento = as.numeric(Rendimiento)) %>% group_by(Nommunicipio,Nommodalidad) %>% summarise(AvgYield = mean(Rendimiento, na.rm=T) *1000, MedYield = median(Rendimiento, na.rm=T) *1000, MinYield = min(Rendimiento, na.rm=T) *1000, MaxYield = max(Rendimiento, na.rm=T) *1000, StDevYield = sd(Rendimiento, na.rm=T) *1000, AvgArea = mean(Cosechada, na.rm=T)) %>% ungroup() %>% mutate(AG_Type = ifelse(Nommodalidad == "Riego", "Irrigation", "Temporal")) %>% rename(Municipio = Nommunicipio) %>% select(-Nommodalidad)

# Join the data with the polygons
AvgYield_sf = HuexModernCultivArea_Municip_AGType %>% left_join(x, by=c("Municipio", "AG_Type"))

# create raster of average yield 
AvgYield = AvgYield_sf %>% select(-AvgArea, -OID_1) %>% #st_rasterize(template=DEM)#, align=T

rm(mun, x, HuexModernCultivArea_Municip_AGType)
```


!!!! st_rasterize(template=DEM)#, align=T --- DOES THIS WORK??




INTENSITY???
--Area of polygon
--Avg Area cultivated
--Avg Area cultivated / Area of polygon = intensity of maize land use??
--pop?
--popdens?
--AG employment? per ha? 2010


NEXT
GAM MODELS - IRRIGATION + TEMPORAL

THEN
EXTRAPOLTE TO COLONIAL

SAME W COCHINEAL

THEN SALIENT AG/ENV VARS TO POLYGONS



# Topographic/Environmental Metrics



## Topographic Variables

DEM_r <- terra::rast(raster(paste0(wd$data_r, "SRTM_30m_DEM_PueblaTlaxcala2_Smooth3.tif")))
DEM = read_stars(paste0(wd$data_r, "SRTM_30m_DEM_PueblaTlaxcala2_Smooth3.tif"))
```{r, label='Topographic Variables', message=FALSE,warning=FALSE}

#create temp directory for rasters we will delete later
# we will stack and save these together at the end of this chunk
dir.create(paste0(wd$data_p,"temp_rasts"))

######### TRI ######### 

TRI <- terra::terrain(DEM_r, v="TRI")# calc Terrain Ruggedness Index
TRI <- st_as_stars(RescaleSpatRast(TRI)) # rescale 0-1


######### TWI ######### 

wbt_breach_depressions(dem = paste0(wd$data_r,"SRTM_30m_DEM_PueblaTlaxcala2_Smooth3.tif"), output = paste0(wd$data_p,"temp_rasts/DEM_breach.tif"))
DEM_breach <- rast(paste0(wd$data_p, "temp_rasts/DEM_breach.tif"))
DEM_breach <- project(DEM_breach,DEM_r)
writeRaster(DEM_breach,paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), overwrite=TRUE)

#DEM2 <- as(DEM, "Raster")
slope <- terra::terrain(DEM_r, v="slope", unit="degrees")#Slope
slope <- project(slope,DEM_r)
writeRaster(slope,paste0(wd$data_p,"temp_rasts/Slope.tif"), overwrite=TRUE)

wbt_d8_flow_accumulation(paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), output=paste0(wd$data_p,"temp_rasts/Accum.tif"), out_type = "specific contributing area")
Accum <- rast(paste0(wd$data_p, "temp_rasts/Accum.tif"))
Accum <- project(Accum,DEM_r)
Accum[is.nan(Accum)] <- NA
writeRaster(Accum,paste0(wd$data_p,"temp_rasts/Accum.tif"), overwrite=TRUE)

wbt_wetness_index(sca=paste0(wd$data_p,"temp_rasts/Accum.tif"), slope=paste0(wd$data_p,"temp_rasts/Slope.tif"), output=paste0(wd$data_p,"temp_rasts/TWI.tif"), verbose_mode = FALSE)
TWI <- rast(paste0(wd$data_p, "temp_rasts/TWI.tif"))
TWI <- project(TWI,DEM_r)
writeRaster(TWI,paste0(wd$data_p,"temp_rasts/TWI.tif"), overwrite=TRUE)

TWI.01 <- RescaleSpatRast(TWI) # rescale 0-1
TWI.12 <- TWI.01+1

######### SPI ######### 

wbt_stream_power_index(sca=paste0(wd$data_p,"temp_rasts/Accum.tif"),slope=paste0(wd$data_p,"temp_rasts/Slope.tif"),output=paste0(wd$data_p,"temp_rasts/SPI.tif"),exponent = 1)
SPI <- rast(paste0(wd$data_p, "temp_rasts/SPI.tif"))
SPI <- project(SPI,DEM_r)
writeRaster(SPI,paste0(wd$data_p,"temp_rasts/SPI.tif"), overwrite=TRUE)
#SPI <- SPI * SaltLakeBinaryInvNA
SPI.01 <- RescaleSpatRast(SPI)
SPI.rs <- RescaleSpatRast(SPI.01)+0.5
#SPI.rs <- SPI.rs * WetAgInv

######### STI ######### 

wbt_fd8_flow_accumulation(paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), output=paste0(wd$data_p,"temp_rasts/Accum_fd8.tif"), out_type = "specific contributing area")
Accum_fd8 <- rast(paste0(wd$data_p, "temp_rasts/Accum_fd8.tif"))
Accum_fd8 <- project(Accum_fd8,DEM_r)
writeRaster(Accum_fd8,paste0(wd$data_p,"temp_rasts/Accum_fd8.tif"), overwrite=TRUE)

wbt_sediment_transport_index(sca=paste0(wd$data_p,"temp_rasts/Accum_fd8.tif"),slope=paste0(wd$data_p,"temp_rasts/Slope.tif"),output=paste0(wd$data_p,"temp_rasts/STI.tif"))
STI <- rast(paste0(wd$data_p, "temp_rasts/STI.tif"))
STI <- project(STI,DEM_r)
writeRaster(STI,paste0(wd$data_p,"temp_rasts/STI.tif"), overwrite=TRUE)
STI.12 <- RescaleSpatRast(STI)
STI.12 <- STI.12+1

######### Curv ######### 

wbt_profile_curvature(dem = paste0(wd$data_p,"temp_rasts/DEM_breach.tif"), output = paste0(wd$data_p,"temp_rasts/Curv.tif"))
Curv <- rast(paste0(wd$data_p, "temp_rasts/Curv.tif"))
Curv <- project(Curv,DEM_r)
writeRaster(Curv,paste0(wd$data_p,"temp_rasts/Curv.tif"), overwrite=TRUE)
Curv[Curv < -0.001] <- -0.001
Curv.rs <- RescaleSpatRast(Curv)*1.5

######### DrainDens Inputs ######### 

Accum2 <- Accum
Accum2[Accum2 > 10000] <- 10000
Accum2<- RescaleSpatRast(Accum2)


```


### Convert to Stars
  
```{r, label='', message=FALSE,warning=FALSE}

```




## Climate Data

!!!! Change resolution and extent of CHELSA and SoilGrids to match DEM


```{r, label='Import CHELSA Climate Data', message=FALSE,warning=FALSE}

MeanPrecip = read_stars(paste0(wd$data_r, "CHELSA_bio10_MeanPrecip.tif"))
MeanTemp = read_stars(paste0(wd$data_r, "CHELSA_bio10_meanTemp.tif"))
NPP = read_stars(paste0(wd$data_r, "CHELSA_NPP.tif"))
GrowSeasLen = read_stars(paste0(wd$data_r, "CHELSA_gsl_GrowingSeasonLength.tif"))
GrowSeasPrecip = read_stars(paste0(wd$data_r, "CHELSA_gsp_GrowingSeasonPrecipSum.tif"))
GrowSeasTemp = read_stars(paste0(wd$data_r, "CHELSA_gst_GrowingSeasonMeanTemp.tif"))

library(spatialEco)
y = as(MeanTemp, "Raster")
x = raster(DEM_r)
MeanTemp.ds <- spatialEco::raster.downscale(x, y, scatter=TRUE, residuals = TRUE,
                            uncertainty = "prediction", se = TRUE)
plot(MeanTemp.ds$downscale, main="Downscaled Mean Temp")
```






## Soil Data

!!!! Change resolution and extent of CHELSA and SoilGrids to match DEM

```{r, label='Import SoilGrids250 Data', message=FALSE,warning=FALSE}

BulkDens = read_stars(paste0(wd$data_r, "SoilGrids250_BulkDensity.tif"))
Clay = read_stars(paste0(wd$data_r, "SoilGrids250_Clay.tif"))
CoarseFrag = read_stars(paste0(wd$data_r, "SoilGrids250_CoarseFragments.tif"))
Nitrogen = read_stars(paste0(wd$data_r, "SoilGrids250_Nitrogen.tif"))
OCDens = read_stars(paste0(wd$data_r, "SoilGrids250_OrganicCarbonDensity.tif"))
Sand = read_stars(paste0(wd$data_r, "SoilGrids250_Sand.tif"))
Silt = read_stars(paste0(wd$data_r, "SoilGrids250_Silt.tif"))
SOC = read_stars(paste0(wd$data_r, "SoilGrids250_SoilOrganaicCarbon.tif"))
SOCStock = read_stars(paste0(wd$data_r, "SoilGrids250_SoilOrganaicCarbonStock.tif"))
WRBSoil = read_stars(paste0(wd$data_r, "SoilGrids250_WRBSoilType.tif"))

```



## Calculate Metrics for Municipal Polygons

Can you do this in STARS????

```{r, "Catchment Area Topographic/Environmental Metrics", warning = FALSE, message=FALSE}

Catch_List2 <- list() #create output list

for(i in 1:length(Catch_List)){
  
  #define catchment area polys as temp object
  tmp.p <- Catch_List[[i]]
  
  #convert to SpatVectors class for use with terra package (much much faster)
  tmp.p2 <- vect(tmp.p)
  
  # NPP
  x <- terra::extract(NPP, tmp.p2, fun=sum, na.rm=T)
  tmp.p@data$NPP.tot <- x[,2]
  tmp.p@data$NPP.avg <- tmp.p@data$NPP.tot/tmp.p@data$Catchment_ha
  
  # Env Zone
  x <- terra::extract(EZ, tmp.p2, fun=mean, na.rm=T)
  tmp.p@data$EZ.avg <- x[,2]
  x <- terra::extract(EZ, tmp.p2, fun=sd, na.rm=T)
  tmp.p@data$EZ.sd <- x[,2]
  
  # TRI (Terrace Need)
  x <- terra::extract(TRI, tmp.p2, fun=sum, na.rm=T)
  tmp.p@data$TRI.tot <- x[,2]
  tmp.p@data$TRI.avg <- tmp.p@data$TRI.tot/tmp.p@data$Catchment_ha
  
  ## Hydraulic Agriculture Potential
  #Irrigation Potential
  x <- terra::extract(TWI.01, tmp.p2, fun=sum, na.rm=T)
  tmp.p@data$IrrigPot.tot <- x[,2]
  tmp.p@data$IrrigPot.avg <- tmp.p@data$IrrigPot.tot/tmp.p@data$Catchment_ha
  #Wetland Agriculture Potential
  x <- terra::extract(WetAgPot, tmp.p2, fun=sum, na.rm=T)
  tmp.p@data$WetAgPot.tot <- x[,2]
  tmp.p@data$WetAgPot.avg <- tmp.p@data$WetAgPot.tot/tmp.p@data$Catchment_ha
  
  # Intensification costs
  x <- terra::extract(IntnsCost, tmp.p2, fun=sum, na.rm=T)
  tmp.p@data$IntnsCost.tot <- x[,2]
  tmp.p@data$IntnsCost.avg <- tmp.p@data$IntnsCost.tot/tmp.p@data$Catchment_ha
  # Intensification costs excluding wetland agriculture
  x <- terra::extract(IntnsCostNW, tmp.p2, fun=sum, na.rm=T)
  tmp.p@data$IntnsCostNW.tot <- x[,2]
  tmp.p@data$IntnsCostNW.avg <- tmp.p@data$IntnsCostNW.tot/tmp.p@data$Catchment_ha
  
  # Agricultural Potential
  x <- terra::extract(AGPot, tmp.p2, fun=sum, na.rm=T)
  tmp.p@data$AGPot.tot <- x[,2]
  tmp.p@data$AGPot.avg <- tmp.p@data$AGPot.tot/tmp.p@data$Catchment_ha
  # Agricultural Potential excluding wetland agriculture
  x <- terra::extract(AGPotNW, tmp.p2, fun=sum, na.rm=T)
  tmp.p@data$AGPotNW.tot <- x[,2]
  tmp.p@data$AGPotNW.avg <- tmp.p@data$AGPotNW.tot/tmp.p@data$Catchment_ha
  
  # Population Pressure
  tmp.p@data$PopPressure <- tmp.p@data$Population.s2/tmp.p@data$AGPot.tot
  tmp.p@data$PopPressureNW <- tmp.p@data$Population.s2/tmp.p@data$AGPotNW.tot
  
  # Erosion Potential
  x <- terra::extract(ErosionPot, tmp.p2, fun=sum, na.rm=T)
  tmp.p@data$ErosionPot.tot <- x[,2]
  tmp.p@data$ErosionPot.avg <- tmp.p@data$ErosionPot.tot/tmp.p@data$Catchment_ha
  
  Catch_List2[[i]] <- tmp.p #save to output list
}

# rename list items
catch.names <- names(Catch_List)
names(Catch_List2) <- catch.names
Catch_List <- Catch_List2
rm(Catch_List2)
```


## Agricultural Intensity


INTENSITY???
--Area of polygon
--Avg Area cultivated
--Avg Area cultivated / Area of polygon = intensity of maize land use??
--pop?
--popdens?
--AG employment? per ha? 2010




## Reorganize and Export Raster Data

!!! DO THIS IN STARS


```{r, "Reorganize and Export Raster Data", warning = FALSE, message=FALSE}
TopoEnvData_RastStack <- c(DEM, Curv, Curv.rs, Accum, Accum_fd8, Accum2, AGPot, AGPotNW, EZ, FreshLakeDepth, FreshLakeDepth.12, IntnsCost, IntnsCostNW, LakesBinaryInv, LakesBinaryInvNA, NPP, NPP.rs, NPP.rs.wetMod, SaltLakeBinaryInv, SaltLakeBinaryInvNA, slope, SPI, SPI.01, SPI.01.wetMod, SPI.rs, STI, STI.12, TRI, TRI.01, TRI.01.wetMod, TWI, TWI.01, TWI.12, TWI.12.wetMod, WetAgInv, WetAgPot, WetAgPot2, ErosionPot)

nam = c("DEM", "Curv", "Curv.rs", "Accum", "Accum_fd8", "Accum2", "AGPot", "AGPotNW", "EZ", "FreshLakeDepth", "FreshLakeDepth.12", "IntnsCost", "IntnsCostNW", "LakesBinaryInv", "LakesBinaryInvNA", "NPP", "NPP.rs", "NPP.rs.wetMod", "SaltLakeBinaryInv", "SaltLakeBinaryInvNA", "slope", "SPI", "SPI.01", "SPI.01.wetMod", "SPI.rs", "STI", "STI.12", "TRI", 'TRI.01', "TRI.01.wetMod", "TWI", "TWI.01", "TWI.12", "TWI.12.wetMod", "WetAgInv", "WetAgPot", "WetAgPot2", "ErosionPot")

names(TopoEnvData_RastStack) = nam

terra::writeRaster(TopoEnvData_RastStack, paste0(wd$data_p, "TopoEnvData_RastLayers.tif"), overwrite=T)
terra::writeRaster(TopoEnvData_RastStack, paste0(wd$data_f, "TopoEnvData_RastLayers.tif"), overwrite=T)

unlink(paste0(wd$data_p, "temp_rasts"), recursive = TRUE)

rm(Catch_List2, x, DEM, Curv.01, Curv, Curv.rs, Accum, Accum_fd8, Accum2, AGPot, AGPotNW, EZ, FreshLakeDepth, FreshLakeDepth.12, IntensCost, IntnsCostNW, LakesBinaryInv, LakesBinaryInvNA, NPP, NPP.01, NPP.rs, NPP.rs.wetMod, NPP2, SaltLakeBinaryInv, SaltLakeBinaryInvNA, slope, SPI, SPI.01, SPI.01.wetMod, SPI.rs, STI, STI.12, tmp.p, tmp.p2, tmp.r, TRI, TRI.01, TRI.01.wetMod, TRI.12, TRI.rs, TWI, TWI.01, TWI.12, TWI.12.wetMod, WetAgInv, WetAgPot, WetAgPot2, ErosionPot, IntnsCost,npp,EZm,TopoEnvData_RastStack)
```



# Modelling Modern Crop Yields

GAM MODELS - IRRIGATION + TEMPORAL

Or Random Forest

Use LASSO routine



## Temporal


## Irrigation



# Estimating 16th Century Yields

EXTRAPOLTE TO COLONIAL

Apply model to Full Arable Area (HuexArable5)

This will now be a raster model, using HuexArable5 as a mask

!!! Arable Area (HuexArable5) needs partition between Irrigation + Temporal
-- Use wetland AG Lopez Corral map to estimate area of irrigation -- in addition to modern distribution of irrigation agriculture

!!! Apply temporal model everywhere first to estimate rainfed yields?? THEN apply irrigation overlay for restricted area???

!!! Do modern yields need rescaling to reach 16th century yields?? Or does the greater labor intensity per unit area + multicropping (companion planting) C16 make the colonial yields comparable to modern ones??
--need to look at colonial recorded maize yields and convert to kg/ha

!!! Each raster cell is how many ha? --> convert each cell to the gross amount of kg maize
--then to people per nutritional reqs --> carrying capacity of each cell

## Temporal


## Irrigation





# Upload Agricultural, Environmental and Topo metrics to MH_Poly














































# Cochineal Production

--greater towards Cholula-Puebla... market forces? Environmental factors?
--N-S gradient in Huex/Salvador
--less in the north and south, more in the east and center
--biased by cabecera! Cholula > Calpan >> Huex/Salvador (Markets)... you can see this clearly comparing Huex and Calpan, and especially the sujetos of Huex vs Calpan in the Atlixco Valley
--Smaller settlements have higher New-Old ratio == production expanding over time in smaller settlements 
--Larger settlements tend to have more Old Nopales per cap (market forces?)... !! Older pop centers (fewer renters + haciendas) tend to be the dominant cultivators of Cochineal
------> this suggests that renters were forced to cultivate estates/haciendas for AG surplus, with much lower market access (contributing to the pattern of RdC-govt-tribute specialization in renter zones), while freeholders nearer to markets were free to exploit the cultivation of grana
--Huex and Atlixco may have lower per cap due to higher agricultural productivity(?)
--greater intensity + land use of cochineal production in more heavily populated areas(!). You would think this would be the inverse -- that lower popdens == less pop pressure == more land/ability to cultivate cochineal. But it looks like the key factors are climate, market access, and polit-econ demands/control of labor.
--This in-turn helps explain why renters in Tepetlaoztoc often had more land than freeholders. Not only did they have to pay rents on this land + be subject to the demands of the landowners, they were encumbered by the labor demands of larger-scale staple agriculture + part-time crafting AT THE EXPENSE OF OTHER MORE LUCRATIVE MARKET GARDENING (turkeys, firewood, tlameme, gathering, cochineal) cornered by the freeholders surrounding population centers

Calculate:
--pop density
--dist to Puebla
--Longer leagues
--Actual distances
--Accessibility pop
--Accessibility cabecera
--Accessibility haciendas

```{r, label='Import Data', message=FALSE,warning=FALSE}

# Cochineal data from 1600
Cochineal1600 = st_read(paste0(wd$data_r,"Cochineal1600.gpkg"))
Cochineal1600h = st_read(paste0(wd$data_r,"Cochineal1600.gpkg")) %>% filter(
  Cabecera == "Huexotzinco" |  Cabecera == "San Salvador") %>% filter(North > 2100000)
```



> plot(Cochineal1600["Cabecera"], pch=19)
> plot(Cochineal1600["OldPerCap"], pch=19)
> plot(Cochineal1600["NewPerCap"], pch=19)
> plot(Cochineal1600["TotalPerCap"], pch=19)

plot(Cochineal1600h["TotalPerCap"], pch=19)
> plot(Cochineal1600h["OldPerCap"], pch=19)
> plot(Cochineal1600h["NewPerCap"], pch=19)
> plot(Cochineal1600h$Vecinos, Cochineal1600h$OldPerCap)
> plot(Cochineal1600h$Vecinos, Cochineal1600h$Dist_Leagues)
> plot(Cochineal1600h$OldPerCap, Cochineal1600h$Dist_Leagues)
> plot(Cochineal1600h$NewPerCap, Cochineal1600h$Dist_Leagues)
> plot(Cochineal1600h$TotalPerCap, Cochineal1600h$Dist_Leagues)
> plot(Cochineal1600h$NewOldRatio, Cochineal1600h$Dist_Leagues)
> plot(Cochineal1600h$NewOldRatio, Cochineal1600h$Vecinos)






