---
title: "Huexotzinco-Tlaxcala C16, Script #X:"
subtitle: "env Data prep"
author: "Rudolf Cesaretti"
date: "Last run on `r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
    number_sections: true
bibliography: References.bib
csl: apa.csl
link-citations: yes
---

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

```{r, setup, include=FALSE,echo=FALSE, message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=75),tidy=TRUE)
#
rm(list = ls())
```

```{r, label='Set Local Directory Location', message=FALSE,warning=FALSE}

wd <- list()

#SET YOUR LOCAL DIRECTORY LOCATION HERE:
#wd$dir <- "C:/Users/rcesaret/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"
wd$dir <- "C:/Users/TJ McMote/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"

wd$analysis <- paste0(wd$dir,"analysis/")
wd$data_r <- paste0(wd$dir,"data-raw/")
wd$data_p <- paste0(wd$dir,"data-processed/")
wd$data_f <- paste0(wd$dir,"data-final-outputs/")
wd$figs <- paste0(wd$dir,"figures/")
wd$funcs <- paste0(wd$dir,"functions/")

```


```{r, label='Load Libraries', message=FALSE,warning=FALSE}
# Package names
packages <- c("rgdal", "rgeos", "sp", "sf", "GISTools", "raster", "Matrix", "terra",
              "tidyterra", "data.table", "cowplot", "scam", "rmarkdown", "spatialreg", 
              "gdistance", "lwgeom", "tidyverse", "tidyr", "stars", "dismo", "spdep",
              "purrr", "spatialEco", "whitebox", "classInt","ggnewscale", "lbmech")#, "data.table", "zoo", "era", "JOPS", "mgcv","igraph",  "ggrepel","ggridges", "movecost",  "datplot", "scales",

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(packages,installed_packages)

#Read in custom R functions located in the wd$funcs directory folder
FUNCS <- list("RescaleSpatRast.R")#"splitByAttributes.R", 
invisible(lapply(FUNCS, function(x) source(paste0(wd$funcs,x))))
rm(FUNCS)

```

setwd("D:/lbmech/Altar") 
dir <- getwd()

# Define Spatial Projection; INEGI Conformal Conic
proj <- "+proj=lcc +lat_0=12 +lon_0=-102 +lat_1=17.5 +lat_2=29.5 +x_0=2500000 +y_0=0"

```{r}

# Import productivity data and locations

prods_maize <- fread(paste0(wd$data_r, "maize_prods.gz"))

locs <- fread(paste0(wd$data_r, "locations.gz"),encoding='UTF-8')

# Calculate variables of interest

prods_maize[, Total_ha := Active_ha + Fallow_ha, # Total area is active + fallow
              ][, LUF := Total_ha / Active_ha # Get amount of land needed per ha active
                ][, Gross := Maize_kg / Maize_ha # Gross amount considers only planted areas
                ][, Prod := Gross / LUF # Total productivity corrects for fallow needs
                  ][, Fert := Fert_ha / Active_ha # Get probability of fertilizer use
                  ][, Irr := MaizeIrr_ha / Maize_ha] # Get probability of irrigation

prods_maize[Size != 'Total'][1:10,.(NOM, Maize_ha, Maize_tons = Maize_kg, Size, LUF,Gross,Prod,Fert,Irr)]

```

The dataset contains up to three observations per municipality—one for large farms above 5 ha, one for small farms below 5 ha, and one for ejidos—a government-enforced agricultural cooperative. There is also a Total entry for each municipality that serves as a checksum (not available for censuses before 1950). Active_ha is the total area planted in any crop; Fallow_ha is the total area fallow in 1960; Maize_ha is the total area planted in maize; MaizeIrr_ha is the total area planted in maize and irrigated; Maize_ton is the tons of maize produced in 1960; Fert_ha is the total area with applied fertilizer.

We need to (1) calculate the yield in kilograms per hectare; (2) account for fallowed fields—Sanders’ ‘land-use factor’; and (3) calculate covariates for percent irrigated and percent fertilized:


list.files(path = "/Users/admin/apps/csv-courses/", pattern = "*.csv")


x = c(
"avhrr-only-v2.19810901.nc",
"avhrr-only-v2.19810902.nc",
"avhrr-only-v2.19810903.nc",
"avhrr-only-v2.19810904.nc",
"avhrr-only-v2.19810905.nc",
"avhrr-only-v2.19810906.nc",
"avhrr-only-v2.19810907.nc",
"avhrr-only-v2.19810908.nc",
"avhrr-only-v2.19810909.nc"
)
# see the second vignette:
# install.packages("starsdata", repos = "http://pebesma.staff.ifgi.de", type = "source")
file_list = system.file(paste0("netcdf/", x), package = "starsdata")
(y = read_stars(file_list, quiet = TRUE))
## stars object with 4 dimensions and 4 attributes
## attribute(s), summary of first 1e+05 cells:















paste0("C:/Users/TJ McMote/Downloads/GFSAD/", x)
CropExtentCMex32614.gpkg
```{r}
Extent4326 <- st_read(paste0(wd$data_r,"CropExtentCMex4326.gpkg"))


x = list.files(path = "C:/Users/TJ McMote/Downloads/GFSAD/GFSAD30NACE/", full.names = TRUE)
y = lapply(x, read_stars)
GFSAD30NACE <- do.call(st_mosaic,y) %>% st_crop(Extent4326) %>% as("Raster")#st_transform(crs=32614)
GFSAD30NACE =  

x = list.files(path = "C:/Users/TJ McMote/Downloads/GFSAD/", full.names = TRUE)

z = lapply(y, st_mosaic)


z=c(y, along=3)



x = list.files(path = "C:/Users/TJ McMote/Downloads/GFSAD/GFSAD30NACE/")
file_list = system.file(paste0("C:/Users/TJ McMote/Downloads/GFSAD/GFSAD30NACE/", x), package = "starsdata")
(y = st_mosaic(x, quiet = TRUE))
```














