---
title: ":"
subtitle: "Roads, Transport and Distances"
author: "Rudolf Cesaretti"
date: "Last run on `r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
    number_sections: true
bibliography: References.bib
csl: apa.csl
link-citations: yes
---


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

```{r, setup, include=FALSE,echo=FALSE, message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=75),tidy=TRUE)
rm(list = ls())
```

```{r, label='Set Local Directory Location', message=FALSE,warning=FALSE}

wd <- list()

#SET YOUR LOCAL DIRECTORY LOCATION HERE:
#wd$dir <- "C:/Users/rcesaret/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"
wd$dir <- "C:/Users/TJ McMote/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"

wd$analysis <- paste0(wd$dir,"analysis/")
wd$data_r <- paste0(wd$dir,"data-raw/")
wd$data_p <- paste0(wd$dir,"data-processed/")
wd$data_f <- paste0(wd$dir,"data-final-outputs/")
wd$figs <- paste0(wd$dir,"figures/")
wd$funcs <- paste0(wd$dir,"functions/")

```


```{r, label='Load Libraries', message=FALSE,warning=FALSE}
# Package names
packages <- c("rgdal", "rgeos", "sp", "sf", "GISTools", "raster", "Matrix", 
              "leastcostpath", "gdistance", "lwgeom", "tidyverse","tidyr", 
              "stars", "dismo", "spatialEco", "whitebox", "classInt", "deldir",
              "spdep", "igraph", "dbscan", "cowplot", "deldir", "cccd", 
              "ggraph", "geosphere", "statnet", "intergraph","ggnewscale")
              
# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(packages,installed_packages)

#Read in custom R functions located in the wd$funcs directory folder
#FUNCS <- list("splitByAttributes.R", "CatchmentPolys.R", "SpatialNets.R", 
#              "net_stats.R", "bbox_polygon.R")
#invisible(lapply(FUNCS, function(x) source(paste0(wd$funcs,x))))
#rm(FUNCS)

```


I do five things in this R markdown document: 

  1. Create site catchment areas
  2. Calculate catchment area transport route densities
  3. Calculate catchment area demographic densities
  4. Define spatial network models for AggSites/Catchments and associated metrics
  5. Reorganize the data and export for Script #5


## Import Data

```{r, label='Import Data', message=FALSE,warning=FALSE}

# Read-in the data
MH_Agg_Poly <- st_read(paste0(wd$data_r,"MH_Poly.gpkg")) %>% rename(AggZone_ID = id, AggZone = Name)
#All_AggPoly <- readOGR(paste0(wd$data_p,"SBOM_AggPoly2.gpkg"))
MH_Agg_Pts = st_centroid(MH_Agg_Poly)


MH_Settlements <- read.csv(paste0(wd$data_r,"MH_Settlements.csv"))

MH_Settlements_Pts <- st_as_sf(MH_Settlements, coords=c("East","North"), crs=32614)



MH_Agg_Roads <- read.csv(paste0(wd$data_r,"HuexRoads.csv"))

# read in 90m DEM and convert the CRS to the same as spatial points dataframe
#DEM <- raster(paste0(wd$data_r, "SBOM_DEM_90m.tif"))
#DEM <- projectRaster(DEM, crs = 26914)

```



# Simple Geodesic Distance Matrices

While the "Transport Route Density" least cost path rasters must be calculated as above, we _can_ use non-LCP geodesic 'great circle' distances instead of LCP cost-distances. The advantage here is that geodesic distance matrices can be calculated in just a few seconds using the code below. 

```{r, label='Simple Geodesic Distance Matrices',  message=FALSE, warning=FALSE}

MH_Settlements_sgDMat = pointDistance(MH_Settlements_Pts, lonlat=F, allpairs=T) / 1000
rownames(MH_Settlements_sgDMat) <- MH_Settlements_Pts$Name
colnames(MH_Settlements_sgDMat) <- MH_Settlements_Pts$Name
write.csv(MH_Settlements_sgDMat, file=paste0(wd$data_p,"MH_Settlements_sgDMat.csv"))


MH_Agg_sgDMat = pointDistance(MH_Agg_Pts, lonlat=F, allpairs=T) / 1000
rownames(MH_Agg_sgDMat) <- MH_Agg_Pts$AggZone
colnames(MH_Agg_sgDMat) <- MH_Agg_Pts$AggZone
write.csv(MH_Agg_sgDMat, file=paste0(wd$data_p,"MH_Agg_sgDMat.csv"))

```


## Simple Geodesic Distance Variables

For AggZones + Settlements
-sgDist_Market = simple geodesic dist to market / dist to regional cabecera
-sgDist_Ciudad = simple geodesic dist to Huex
-sgDist_Pueblo = simple geodesic dist to pueblo core

```{r, label='Simple Geodesic Distance Variables',  message=FALSE, warning=FALSE}

```








# Road Network Geodesic Distance Matrix for AggZones

```{r, label='Road Network Geodesic Distance Matrix for AggZones',  message=FALSE, warning=FALSE}

g = graph_from_edgelist(cbind(MH_Agg_Roads$Origin, MH_Agg_Roads$Destination))
E(g)$weight = MH_Agg_Roads$Dist_km
MH_Agg_rDMat = distances(g)
rm(g)

MH_Agg_rDMat <- MH_Agg_rDMat[rownames(MH_Agg_sgDMat),,drop=FALSE]
MH_Agg_rDMat <- MH_Agg_rDMat[,colnames(MH_Agg_sgDMat),drop=FALSE]

write.csv(MH_Agg_rDMat, file=paste0(wd$data_p,"MH_Agg_rDMat.csv"))

```


## Road Network Geodesic Distance Variables for AggZones

-rDist_Market = road network geodesic dist to market / dist to regional cabecera
-rDist_Ciudad = road network geodesic dist to Huex
-rDist_Pueblo = road network geodesic dist to pueblo core

```{r, label='Road Network Geodesic Distance Variables for AggZones',  message=FALSE, warning=FALSE}

```










































































































