---
title: "Sectors & Industries"
author: "Rudolf Cesaretti"
date: "2023-03-05"
output: html_document
---


```{r, setup, include=FALSE,echo=FALSE, message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=75),tidy=TRUE)
#
rm(list = ls())
```



# Setup 

All of the data and scripts are downloadable from the [new ASU SettlementPersist2022 github repository](https://https://github.com/rcesaret/ASUSettlementPersist2022), which can be downloaded locally as a .zip folder or cloned to your own account.

Either way, once you have done so, you will need to modify the working directory (setwd("C:/...)") path and "dir" variables in the code chunk below to match the repository location on your computer.

```{r, label='Set Local Directory Location', message=FALSE,warning=FALSE}

wd <- list()

#SET YOUR LOCAL DIRECTORY LOCATION HERE:
#wd$dir <- "C:/Users/rcesaret/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"
wd$dir <- "C:/Users/TJ McMote/Dropbox (ASU)/Huexotzinco/HuexTlaxAnal2022/HuexTlax2022Analysis/"

wd$analysis <- paste0(wd$dir,"analysis/")
wd$data_r <- paste0(wd$dir,"data-raw/")
wd$data_p <- paste0(wd$dir,"data-processed/")
wd$data_f <- paste0(wd$dir,"data-final-outputs/")
wd$figs <- paste0(wd$dir,"figures/")
wd$funcs <- paste0(wd$dir,"functions/")

```


## Load R Packages and Custom Functions

```{r, label='Load Libraries', message=FALSE,warning=FALSE}
# Package names
packages <- c("rgdal", "sp", "sf", "GISTools", "lwgeom", "tidyverse", "tidyr", 
              "data.table", "zoo", "scales", "sjmisc")
              
              #, "data.table",  "mgcv","igraph", "ggrepel","ggridges", "movecost",  "datplot", "scales",

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(packages,installed_packages)

#Read in custom R functions located in the wd$funcs directory folder
#FUNCS <- list("splitByAttributes.R", "RS_Acoef.R", "PopAreaResids.R", "LQ.R", "GiniSp.R")
#invisible(lapply(FUNCS, function(x) source(paste0(wd$funcs,x))))
#rm(FUNCS)

```


```{r, label='Import Data', message=FALSE,warning=FALSE}

# Read-in the data
MH_Settlements <- read.csv(paste0(wd$data_r,"MH_Settlements.csv"))
MH_Occu <- read.csv(paste0(wd$data_r,"MH_Occu.csv"))
MH_Attr <- read.csv(paste0(wd$data_r,"MH_Attr.csv"), na.strings	= "")

#Poly <- readOGR(paste0(wd$data_r,"SBOMPoly.gpkg"))

```


```{r}

#MH_SD <- left_join(MH_Settlements, MH_Data, by=c('ORD'='ORD', 'Name'='Name'))

MH_Occu_long <- MH_Occu %>% pivot_longer(cols = OldFarm_T:CantSeeIfOccu_U,
  names_to = "Variable", values_to = "Count") %>% filter(Count != 0) %>% 
  left_join(MH_Attr, by="Variable")

MH_Occu_long_Totals <- MH_Occu_long %>% filter(Status == "Total")

MH_Occu_long_NoTotals <- MH_Occu_long %>% filter(Status != "Total")

### Add total columns for every variable??? (Nobles, infirm demog, etc.)


```

# Counts for Sectors, Subsectors and Industries

IMPORTANT: These need STATUSES (T F R U)

```{r}

## Counts for Sectors3

Sect3_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Sectors3) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3)) %>% pivot_wider(id_cols=Name, names_from = Sectors3, names_prefix = "Sect3_Tot_", values_from = Count, values_fill = 0)

Sect3_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors3) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3)) %>% pivot_wider(id_cols=Name, names_from = Sectors3, names_prefix = "Sect3_Tot_NoPix_", values_from = Count, values_fill = 0)

Sect3_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Sectors3) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3)) %>% pivot_wider(id_cols=Name, names_from = Sectors3, names_prefix = "Sect3_Trib_", values_from = Count, values_fill = 0)

Sect3_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors3) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3)) %>% pivot_wider(id_cols=Name, names_from = Sectors3, names_prefix = "Sect3_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Sectors4

Sect4_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Sectors4) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4)) %>% pivot_wider(id_cols=Name, names_from = Sectors4, names_prefix = "Sect4_Tot_", values_from = Count, values_fill = 0)

Sect4_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors4) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4)) %>% pivot_wider(id_cols=Name, names_from = Sectors4, names_prefix = "Sect4_Tot_NoPix_", values_from = Count, values_fill = 0)

Sect4_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Sectors4) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4)) %>% pivot_wider(id_cols=Name, names_from = Sectors4, names_prefix = "Sect4_Trib_", values_from = Count, values_fill = 0)

Sect4_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors4) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4)) %>% pivot_wider(id_cols=Name, names_from = Sectors4, names_prefix = "Sect4_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Sectors3b

Sect3b_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Sectors3b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3b)) %>% pivot_wider(id_cols=Name, names_from = Sectors3b, names_prefix = "Sect3b_Tot_", values_from = Count, values_fill = 0)

Sect3b_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors3b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3b)) %>% pivot_wider(id_cols=Name, names_from = Sectors3b, names_prefix = "Sect3b_Tot_NoPix_", values_from = Count, values_fill = 0)

Sect3b_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Sectors3b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3b)) %>% pivot_wider(id_cols=Name, names_from = Sectors3b, names_prefix = "Sect3b_Trib_", values_from = Count, values_fill = 0)

Sect3b_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors3b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors3b)) %>% pivot_wider(id_cols=Name, names_from = Sectors3b, names_prefix = "Sect3b_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Sectors4b

Sect4b_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Sectors4b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4b)) %>% pivot_wider(id_cols=Name, names_from = Sectors4b, names_prefix = "Sect4b_Tot_", values_from = Count, values_fill = 0)

Sect4b_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors4b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4b)) %>% pivot_wider(id_cols=Name, names_from = Sectors4b, names_prefix = "Sect4b_Tot_NoPix_", values_from = Count, values_fill = 0)

Sect4b_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Sectors4b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4b)) %>% pivot_wider(id_cols=Name, names_from = Sectors4b, names_prefix = "Sect4b_Trib_", values_from = Count, values_fill = 0)

Sect4b_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Sectors4b) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Sectors4b)) %>% pivot_wider(id_cols=Name, names_from = Sectors4b, names_prefix = "Sect4b_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Subsectors

SubSect_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Subsector) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Subsector)) %>% pivot_wider(id_cols=Name, names_from = Subsector, names_prefix = "SubSect_Tot_", values_from = Count, values_fill = 0)

SubSect_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Subsector) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Subsector)) %>% pivot_wider(id_cols=Name, names_from = Subsector, names_prefix = "SubSect_Tot_NoPix_", values_from = Count, values_fill = 0)

SubSect_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Subsector) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Subsector)) %>% pivot_wider(id_cols=Name, names_from = Subsector, names_prefix = "SubSect_Trib_", values_from = Count, values_fill = 0)

SubSect_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Subsector) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Subsector)) %>% pivot_wider(id_cols=Name, names_from = Subsector, names_prefix = "SubSect_Trib_NoPix_", values_from = Count, values_fill = 0)

## Counts for Industries

Industry_Tot = MH_Occu_long_NoTotals %>% group_by(Name, Industry) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Industry)) %>% pivot_wider(id_cols=Name, names_from = Industry, names_prefix = "Industry_Tot_", values_from = Count, values_fill = 0)

Industry_Tot_NoPix = MH_Occu_long_NoTotals %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Industry) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Industry)) %>% pivot_wider(id_cols=Name, names_from = Industry, names_prefix = "Industry_Tot_NoPix_", values_from = Count, values_fill = 0)

Industry_Trib = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% group_by(Name, Industry) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Industry)) %>% pivot_wider(id_cols=Name, names_from = Industry, names_prefix = "Industry_Trib_", values_from = Count, values_fill = 0)

Industry_Trib_NoPix = MH_Occu_long_NoTotals %>% filter(Trib == "Tributary") %>% filter(Occupation != "CenPix") %>% filter(Occupation != "MacPix") %>% group_by(Name, Industry) %>% summarise(Count = sum(Count)) %>% ungroup() %>% filter(!is.na(Industry)) %>% pivot_wider(id_cols=Name, names_from = Industry, names_prefix = "Industry_Trib_NoPix_", values_from = Count, values_fill = 0)


```


```{r}

SectInd_Tot


x = Sect3_Tot %>% left_join(Sect4_Tot, by = "Name") %>% left_join(Sect3b_Tot, by = "Name") %>%
  left_join(Sect4b_Tot, by = "Name") %>% left_join(SubSect_Tot, by = "Name") %>%
  left_join(Industry_Tot, by = "Name") %>% left_join(Sect3_Tot_NoPix, by = "Name") %>%
  left_join(Sect4_Tot_NoPix, by = "Name") %>% left_join(Sect3b_Tot_NoPix, by = "Name") %>%
  left_join(Sect4b_Tot_NoPix, by = "Name") %>% left_join(SubSect_Tot_NoPix, by = "Name") %>%
  left_join(Industry_Tot_NoPix, by = "Name") %>% 
  left_join(Sect3_Trib, by = "Name") %>% left_join(Sect4_Trib, by = "Name") %>%
  left_join(Sect3b_Trib, by = "Name") %>% left_join(Sect4b_Trib, by = "Name") %>%
  left_join(SubSect_Trib, by = "Name") %>% left_join(Industry_Trib, by = "Name") %>%
  left_join(Sect3_Trib_NoPix, by = "Name") %>% left_join(Sect3b_Trib_NoPix, by = "Name") %>%
  left_join(Sect4_Trib_NoPix, by = "Name") %>% left_join(Sect4b_Trib_NoPix, by = "Name") %>%
  left_join(SubSect_Trib_NoPix, by = "Name") %>% left_join(Industry_Trib_NoPix, by = "Name") #%>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name") %>%
  #left_join(XXXXX, by = "Name") %>% left_join(XXXXX, by = "Name")

#x[is.na(x)] <- 0

MH_Occu <- MH_Occu %>% left_join(x, by = "Name") 

```
  
  
```{r}
rm(Sect3_Tot, Sect3_Tot_NoPix, Sect3_Trib, Sect3_Trib_NoPix, Sect4_Tot, Sect4_Tot_NoPix, Sect4_Trib, Sect4_Trib_NoPix, 
Sect3b_Tot, Sect3b_Tot_NoPix, Sect3b_Trib, Sect3b_Trib_NoPix, Sect4b_Tot, Sect4b_Tot_NoPix, Sect4b_Trib, Sect4b_Trib_NoPix, SubSect_Tot, SubSect_Tot_NoPix, SubSect_Trib, SubSect_Trib_NoPix, Industry_Tot, Industry_Tot_NoPix, Industry_Trib, Industry_Trib_NoPix, x)
```

	GovtChurchDemand	SpanishDemand	NobleDemand	LuxuryDemand	UtilDemand